Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-23 csf2/main.tex
nit
change fig3 and hyperref

%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\newcommand{\url}[1]{\lstinline{#1}}

@@

  % \MKin{context}
  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  % \MKin{need}
  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness
  % \MKin{task}
  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time. %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
  % \MKin{object+findings}
  % This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
  % \MKin{object}
  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
  % \MKin{conclusion}
  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
  % Some results of this work are implemented in the Coq proof assistant.

@@

% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
% Programs that attain \gls*{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

@@

This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{mssscct}.
Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
%
In summary, this paper makes the following contributions:
\begin{itemize}
  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
        % 
        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
        % 
        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
        %
        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.

@@

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
The compositionality framework (\Cref{sec:compcomp}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, property satisfaction is defined as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

\bul{A cross-language relation $\sim$ is well-formed with respect to a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
\begin{definition}[Well-formedness of Trace Relations for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
  % $$

@@

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\sigma$]\label{thm:rtpsim:sig}
  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

@@

\Cref{corr:rtp} provides an easy to use theorem to compose secure compilers without any caveats. 
However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed with respect to $\trg{\class[_2]}$, if $\sim_2$ is not well-formed with respect to $\trg{\class[_1]}$, care must be taken. 
Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
\Cref{sec:formalities:compoex} provides an example of this phenomenon.

%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
% 
% \MPin{
% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
%  e.g., remove all deallocations to "optimise" memory management by offloading
%  the work to the OS.
% }
% }

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
\begin{definition}[Upper Composition]
  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
  % Given \texttt{p}, yield
  \lambda \texttt{p}\ldotp
  \left\{\mbox{\begin{tabular}{c}
    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

% Examples of this are present in industry:
% Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
% Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
% Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
% Both languages can be used at the same time in one project~\cite{androidstudio}.
% A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
Now, the following theorem tells us what happens if these are secure:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
\end{theorem}

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
\begin{definition}[Lower Composition]
  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
  % Given $L$ and $\src{p}$, yield
  \lambda \src{p}, L\ldotp
  \left\{\mbox{\begin{tabular}{c}
    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

% The following theorem demonstrates what happens if the involved compilers are secure:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
\end{theorem}

@@

\begin{theorem}[$\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
  $\rtp{\cca}{\tmssafe}$ % \Coqed
\end{theorem}

@@

\begin{theorem}[$\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
  $\rtp{\ccb}{\smssafe}$ % \Coqed
\end{theorem}

\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.

>>>>

\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\newcommand{\url}[1]{\lstinline{#1}}

@@

  % \MKin{context}
  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  % \MKin{need}
  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness
  % \MKin{task}
  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving w.r.t. cryptographic constant time. %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
  % \MKin{object+findings}
  % This paper answers that question positively: The composition of compilers that are secure w.r.t. some security-properties yields the intersection of these properties.
  % \MKin{object}
  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
  % \MKin{conclusion}
  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure w.r.t. both memory safety and cryptographic constant time.
  % Some results of this work are implemented in the Coq proof assistant.

@@

% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
% Programs that attain \gls*{cct} are secure w.r.t. timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

@@

This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{mssscct}.
Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
%
In summary, this paper makes the following contributions:
\begin{itemize}
  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
        % 
        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
        % 
        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
        %
        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.

@@

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
The compositionality framework (\Cref{sec:compcomp}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
This program behaviour is insecure w.r.t. a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, property satisfaction is defined as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

\bul{A cross-language relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
\begin{definition}[Well-formedness of Trace Relations for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
  % $$

@@

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t. class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

@@

\Cref{corr:rtp} provides an easy to use theorem to compose secure compilers without any caveats. 
However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
\Cref{sec:formalities:compoex} provides an example of this phenomenon.

%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
% 
% \MPin{
% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
%  e.g., remove all deallocations to "optimise" memory management by offloading
%  the work to the OS.
% }
% }

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
\begin{definition}[Upper Composition]
  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
  % Given \texttt{p}, yield
  \lambda \texttt{p}\ldotp
  \left\{\mbox{\begin{tabular}{c}
    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

% Examples of this are present in industry:
% Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
% Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
% Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
% Both languages can be used at the same time in one project~\cite{androidstudio}.
% A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
Now, the following theorem tells us what happens if these are secure:
Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
\end{theorem}

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
\begin{definition}[Lower Composition]
  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
  % Given $L$ and $\src{p}$, yield
  \lambda \src{p}, L\ldotp
  \left\{\mbox{\begin{tabular}{c}
    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
  \end{tabular}}\right.$$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

% The following theorem demonstrates what happens if the involved compilers are secure:
Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
\end{theorem}

@@

\begin{theorem}[$\cca$ is secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
  $\rtp{\cca}{\tmssafe}$ % \Coqed
\end{theorem}

@@

\begin{theorem}[$\ccb$ is secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
  $\rtp{\ccb}{\smssafe}$ % \Coqed
\end{theorem}

\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.

\begin{theorem}[$\cca\circ\ccb$ is secure w.r.t. \gls*{ms}]\label{thm:ccab:rtp:ms}
  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
\end{theorem}
% \begin{proof}\MP{keep?}
%   From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
%   Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
%   Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
%   From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
% \end{proof}
